---
title: "Data_vis_final_project"
output: html_document
date: "2026-05-05"
---

Packages 
```{r}
library(ggplot2)
library(scales)
library(tidyverse)
library(MASS)
```


### Accessibility: What is the total amount of active players and how is that split among the two editions?
```{r}
player_base_basic <- data.frame(Edition = c("Java", "Bedrock"), players = c(32000000, 118000000))
player_base <- read.csv("/Users/gwuys/Desktop/School/Data Visualization/Final Project/Data/Player_base.csv", header = T)

### Pie chart of the active player count within each edition
ggplot(player_base_basic, aes(x = "", y = players, fill = Edition)) +
  geom_bar(stat = "identity", width = 1) +
  geom_text(aes(label = label_number(scale = 1e-6, suffix = "M")(players)), 
            position = position_stack(vjust = 0.5)) + 
  scale_fill_manual(values = c("#B3EE3A", "#FFB90F")) +
  coord_polar("y", start = 0) + 
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

### Bar graph of active player count based on electronic system used
ggplot(player_base, aes(x = Edition,y = Player.count, fill = System)) +
  geom_col() +
  geom_text(aes(label = paste0(Player.count/1e6, "M")),
            position = position_stack(vjust = 0.5), size = 3) +
  coord_flip() +
  scale_fill_manual(values = c("#698B22", "#A2CD5A","#B3EE3A", "#FFB90F")) +
  scale_y_continuous(labels = function(x) paste0(x/1e6, "M")) +
  ylab("Player count") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

### Answer: Bedrock wins in regards to accessibility because of its larger selection of eletronic systems that allows for cross-platform gameplay.
```


### Stability: Which edition seems to be more stable based on the amount of bugs that get patched per year? Which years have a significant amount of bugs between both editions?
```{r}
bugs <- read.csv("/Users/gwuys/Desktop/School/Data Visualization/Final Project/Data/Bugs.csv")

### Combining total amount of bugs per year
bugs_yearly <- bugs %>%
  group_by(Year, Edition) %>%
  summarise(total_bugs = sum(Bugs.fixed, na.rm = TRUE)) %>%
  ungroup()


### Ploting the line graph of the total number of bugs fixed per year
ggplot(bugs_yearly, aes(x = Year, y = total_bugs, color = Edition)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  scale_color_manual(values = c("#B3EE3A", "#FFB90F")) +
  labs(
    x = "Year",
    y = "Total bugs fixed"
  ) +
  theme_classic(base_size = 14) + 
  theme(plot.title = element_text(hjust = 0.5))


### Running a Wilcoxon rank sum test to see the difference in the median of total bugs fixed between editions per year
results <- bugs %>%
  group_by(Year) %>%
  summarise(p_value = wilcox.test(Bugs.fixed ~ Edition)$p.value)

results

### Answer: Overall, Java has had lower amounts of bug fixes throughout the years which indicates that the edition is more stable than it's counterpart. The p-values of the Wicoxon rank sum showcase that the notable years with significant differences between the amount of bugs fixed in two editions include 2012, 2018, 2024, and 2025. 
```

### Performance: How fast can players beat the game (defeat the ender dragon)? 
```{r}
speed <- read.csv("/Users/gwuys/Desktop/School/Data Visualization/Final Project/Data/speedrun.csv")

### Boxplot of the fastest times to beat the game across both editions
ggplot(speed, aes(x = Edition, y = IGT.in.mins, fill = Edition)) +
  geom_boxplot(alpha = 0.6) +
  labs(
    x = "Edition",
    y = "IGT (minutes)"
  ) +
  scale_fill_manual(values = c("#B3EE3A", "#FFB90F")) +
 ylab("Time (minutes)") + 
  theme_classic(base_size = 14) + 
  theme(plot.title = element_text(hjust = 0.5), size = 5) + 
  theme(legend.position = "none")

### Running a Wilcoxon rank sum test to see if the median speedrun time in the Bedrock is statistically different from the median speed run time in Java
wilcox.test(IGT.in.mins ~ Edition, data = speed)

### Answer: Java players spent a shorter amount of time completing the game which indicates that they are not experiencing performance issues that would cause their runs to go for longer amounts of time. 
```

### Player Experience: Is there a difference between the average time spent on Minecraft from when the game is launched to when the player disconnects between both editions?
```{r}
sessions <- read_csv("/Users/gwuys/Desktop/School/Data Visualization/Final Project/Data/Session_stats.csv")

### Bar graph of the average daily player session length per edition based on their game experience
ggplot(sessions, aes(x = `player type`,
                     y = average.session.in.mins,
                     fill = Edition)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#B3EE3A", "#FFB90F")) +
  labs(
    x = "Player type",
    y = "Average session length (mins)"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  theme(plot.title = element_text(hjust = 0.5), size = 5)

### Answer: While looking at the plot, it looks to be that Java players have higher average session length across the player types, it is not statistically significant.
```

### Community: Are players more active on servers in one edition over the other? Is there a correlation between active server player count and edition?
```{r}
servers <- read.csv("/Users/gwuys/Desktop/School/Data Visualization/Final Project/Data/servers.csv")

### Making edition categorical
servers$Edition <- factor(servers$Edition)

ggplot(servers, aes(x = Edition, y = Active.Players, fill = Edition)) +
  geom_boxplot() +
  scale_fill_manual(values = c("#B3EE3A", "#FFB90F")) +
  scale_y_log10() + 
  ylab("Active players") +
  theme_classic() + 
  theme(legend.position = "none")

### Running a glm to see whether edition is a significant predictor of server active player count
model <- glm.nb(`Active.Players` ~ Edition, data = servers)
summary(model)

### Categorize servers into 4 sizes based on their active player count
servers <- servers %>%
  mutate(Size = case_when(
    `Active.Players` > 10000 ~ "Mega",
    `Active.Players` >= 1000 ~ "Large",
    `Active.Players` >= 100 ~ "Medium",
    TRUE ~ "Small"
  ))

names(servers)

### Plot box of the amount of servers in 4 categorical sizes according to the active player count
servers %>%
  count(Size, Edition) %>%
  ggplot(aes(
    x = reorder(Size, -n),
    y = n,
    fill = Edition
  )) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#B3EE3A", "#FFB90F")) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) + 
  xlab("Size") + 
  ylab("Number of servers")

### Answer: Java looks to average a higher active server player count but it is not statistically different from Bedrock. 
```
